Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-14 paper/paper.tex
minor

\begin{proof}
We construct an instance of the problem with $\ell$ clusters 
$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
Fix any online algorithm \ALG.
We construct the input sequence of requests for \ALG{} as follows.
First,
we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
nodes on the cluster $S_1$.
Recall that internal requests does not incur any cost for \ALG.
If \ALG{} at any point splits a component
(i.e., spreads its nodes over two or more clusters),
then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
If \ALG{} never collocates them then it is not competitive.

Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
Next,
we issue the request between $x_0$ and $y_0$.
Since this is an external request,
ALG joins them into one component and collocates them in some cluster other than $S_1$.
For this,
ALG moves to a new configuration (i.e., partition)
that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.

Next,
we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
consequently replaces $x_1$ with some other (single) node $x_2$.
We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
 until there are only two single nodes left that  originate from the same cluster.
I.e.,
two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
$x'$ and $y'$,
we have $I(x') \neq I(y')$.
At this point there are at most $\ell+1$ single nodes left,
otherwise there would be more pairs of single nodes that were initially in the same cluster.

Given this sequence of requests,
the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
 swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
Hence,
OPT pays $4$ for node migrations and
ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
\end{proof}

>>>>

\begin{proof}
We construct an instance of the problem with $\ell$ clusters 
$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
Fix any online algorithm \ALG{}.
We construct the input sequence of requests for \ALG{} as follows.
First,
we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
nodes on the cluster $S_1$.
Recall that internal requests does not incur any cost for \ALG{}.
If \ALG{} at any point splits a component
(i.e., spreads its nodes over two or more clusters),
then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
If \ALG{} never collocates them then it is not competitive.

Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
Next,
we issue the request between $x_0$ and $y_0$.
Since this is an external request,
\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
For this,
\ALG{} moves to a new configuration (i.e., partition)
that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.

Next,
we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
consequently replaces $x_1$ with some other (single) node $x_2$.
We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
 until there are only two single nodes left that  originate from the same cluster.
I.e.,
two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
$x'$ and $y'$,
we have $I(x') \neq I(y')$.
At this point there are at most $\ell+1$ single nodes left,
otherwise there would be more pairs of single nodes that were initially in the same cluster.

Given this sequence of requests,
the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
 swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
Hence,
OPT pays $4$ for node migrations and
\ALG{} pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
\end{proof}
